---
title: "A Study of New York Eviction Events Based on Spatial Data Analysis"
author: "Name: Zhu Jinzhou"
student number: 24080728
output: html_document
---
# Originality declaration  

I, [Zhu Jinzhou], confirm that the work presented in this assessment is my own. Where information has been derived from other sources, I confirm that this has been indicated in the work.

Introduction
The aim of this analysis is to explore the spatial distribution characteristics of eviction events that occurred in a specific region (assumed to be a city) in 2020 through spatial data analysis. The primary data sources used include eviction event data (with coordinate information) and community boundary data (in Shapefile format). By applying a series of spatial analysis methods, particularly spatial clustering (DBSCAN) and spatial autocorrelation analysis (Local Moran's I), we aim to understand the distribution patterns of eviction events in different community areas and whether there is any spatial clustering phenomenon.

# Start your response here
```{r}
library(readr)
library(ggplot2)
library(sf)
library(dplyr)
library(spdep)
library(spatstat)
library(sp)
library(tmap)
library(tmaptools)
library(janitor)
library(stringr)
```

```{r}
eiction_points <- read_csv("/Users/troy/Desktop/UCL/PracticeExam/Evictions.csv",na=c(" "))

communty_areas <- st_read("/Users/troy/Desktop/UCL/PracticeExam/Community Districts/geo_export_634c97af-c097-4829-82db-36e4f9d2b899.shp")
```
To ensure that the data has the same projection coordinate system, I transformed all spatial data into the same coordinate reference system (CRS). I then used the tmap package to visualize eviction events and community areas. The graphic shows the distribution of residential eviction events in 2020 by region.
```{r}
points <- eiction_points %>%
  filter(Longitude<0 & Latitude>0)%>%
  st_as_sf(.,coords = c("Longitude","Latitude"),
           crs=4326)
```

```{r}
tmap_mode("plot")
tm_shape(communty_areas)+
  tm_polygons(col=NA,alpha = 0.5)+
tm_shape(points)+
  tm_dots(col="blue")
```

```{r}
communty_areas <- communty_areas%>%
  st_transform(.,4326)
points_sub<- points[communty_areas, ] #,st_intersects()
```

```{r}
points_sub_2020 <- points_sub %>% #Screening point is within the zone
  clean_names() %>%
  filter(str_detect(executed_date, "2020")) %>%
  filter(residential_commercial == "Residential")
```

```{r}
tmap_mode("plot")
tm_shape(communty_areas)+
  tm_polygons(col=NA,alpha = 0.5)+
tm_shape(points_sub_2020)+
  tm_dots(col="blue")
```

```{r}
communty_areas_projected <- communty_areas %>%
  st_transform(.,6538)
points_sub_2020_projected <- points_sub_2020 %>%
  st_transform(.,6538)

window <- as.owin(communty_areas_projected) 
plot(window)

# create a sp object 
points_sub_2020_projected_sp <- points_sub_2020_projected %>%
  as(.,"Spatial")
#create a ppp object
points_sub_2020_projected_sp.ppp <-
ppp(x=points_sub_2020_projected_sp@coords[,1],
    y=points_sub_2020_projected_sp@coords[,2],
    window =window)
```
I use Ripley's K to explore whether eviction events clustered in certain areas. I don't want to use quadrat analysis because it has too many limitations.
```{r}
#Plot Ripley’s K function to analyze spatial clustering or uniform distribution of point patterns
K <- points_sub_2020_projected_sp.ppp %>% 
  Kest(.,correction="border") %>%
  plot()
```
This shows that this incident has obvious clustering characteristics.

Ripley's K analyses are useful exploratory techniques that can tell us if there is spatial clustering in our point data, but they cannot tell us where the clustering occurs in the area we are interested in. That is why I need to use DBSCAN. This method helps to find densely clustered areas and effectively identify noise (i.e. points that do not belong to any cluster).
```{r}
#Density-based spatial clustering of applications with noise: DBSCAN
library(sp)
library(dbscan)
library(fpc)

#first extract the points from the spatial data frame
points_todf <- points_sub_2020_projected_sp %>%
coordinates(.) %>%
as.data.frame()

#The core parameter eps of DBSCAN is set to 1000 meters and MinPts is set to 50. This means that a cluster requires at least 50 points and the maximum distance between adjacent points is 1000 meters.
#now run the dbscan analysis
points_todf_DBSCAN <- points_todf %>%
fpc::dbscan(.,eps=1000,MinPts=50)

points_todf %>%
dbscan::kNNdistplot(.,k=50)
 
plot(points_todf_DBSCAN,points_todf,main="DBSCAN output",frame=F)
plot(communty_areas_projected$geometry,add=T)
```

```{r}
# Get DBSCAN cluster labels.
points_todf$dbcluster <- points_todf_DBSCAN$cluster

# Filter out points with dbcluster > 0 (i.e. non-noise points).
tosf <- points_todf %>%
  st_as_sf(., coords = c("coords.x1", "coords.x2"), crs = 6538) %>%
  filter(dbcluster > 0)
```

```{r}
ggplot(data = communty_areas_projected)+
  geom_sf()+
  geom_sf(data=tosf,size= 0.4,color=tosf$dbcluster,
          fill=tosf$dbcluster)
```

```{r}
check_example <- communty_areas_projected%>%
  st_join(tosf)

points_sf_joined <- communty_areas_projected%>%
  mutate(n=lengths(st_intersects(.,tosf)))%>%
  clean_names()%>%
  mutate(area=st_area(.)) %>%
  mutate(density=n/area)
```

```{r}
tm_shape(points_sf_joined)+
  tm_polygons("density",
              style="jenks",
              palette="PuOr",
              title="Eviction density")
```
As can be seen from the figure, there are obvious purple patches in uptown New York City, so I use Moran's I to check it out.

To further test whether the spatial distribution of evictions is autocorrelated, I use the Local Moran's I method. This method helps identify clustered patterns of evictions within a region (i.e., "high-high" or "low-low" areas).
I use the Queen adjacency rule (i.e., regions that share borders or corners are considered adjacent regions) to create an adjacency matrix that describes the spatial associations between community regions.
```{r}
library(spdep)
coordsw <-points_sf_joined %>%
  st_centroid()%>%
  st_geometry()
plot(coordsw,axes=TRUE)

#make neighbours

community_nb <- points_sf_joined %>%
  poly2nb(.,queen=T)
summary(community_nb)

plot(community_nb,st_geometry(coordsw),col="red")
plot(points_sf_joined$geometry,add=T)
```
As can be seen from the figure, the points are mainly concentrated in the 300000-320000 area.

```{r}
#make weight matrix
community_nb.lw <- community_nb %>%
  nb2mat(.,style ="W")

sum(community_nb.lw)
```

```{r}
community_nb.lw <- nb2listw(community_nb, zero.policy = TRUE)
```

```{r}
I_LWard_Global_Density <- points_sf_joined %>%
  pull(density)%>%
  as.vector()%>%
  localmoran(.,community_nb.lw)
I_LWard_Global_Density
```

The results of Local Moran's I show that some areas have significant high-density (high-high) clustering, while other areas show low-density (low-low) phenomena. In particular, in some high-high areas, the density of eviction events is significantly higher than that in neighboring areas.
```{r}
# Assume I_LWard_Global_Density is a vector
points_sf_joined <- points_sf_joined %>%
  mutate(density_I = as.numeric(I_LWard_Global_Density[,'Ii'])) %>%  # Access the first element using index
  mutate(density_Iz = as.numeric(I_LWard_Global_Density[, 'Z.Ii']))  # Combine the values ​​of Z and Ii.
```
```{r}
breaks1 <-c(-100,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)

library(RColorBrewer)
MoranColours<- rev(brewer.pal(8, "RdGy"))

tm_shape(points_sf_joined) +
    tm_polygons("density_Iz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA,
        title="Local Moran's I, Evictions in New York")
```
In the picture, evictions are more concentrated in the darker areas of Uptown New York than in other areas.

# Results and discussion
From the analysis results, I can draw the following key conclusions:1) Spatial Clustering of Eviction Events: Through DBSCAN clustering analysis and Local Moran's I test, I found a clear spatial clustering of eviction events in certain community areas in 2020. In particular, eviction events occurred at higher rates in certain high-density clustered areas. 2) High-High and Low-Low Areas: The Local Moran's I analysis revealed the spatial distribution characteristics of high-density eviction areas (high-high) and low-density eviction areas (low-low). This provides policymakers with clues about areas that may require more attention. 3) Regional Density Differences: The differences in eviction event density across various community areas provide data support for further research and understanding of the socioeconomic context of eviction events.

# Conclusion
This analysis provides an in-depth exploration of the spatial distribution of eviction events in 2020 through spatial clustering and spatial autocorrelation methods. The results indicate that eviction events exhibit significant spatial clustering in certain community areas, with considerable density variations. These findings offer valuable insights for policy-making and social intervention in related fields. Future research could further investigate the socioeconomic factors of eviction events to gain a more comprehensive understanding of the underlying causes of eviction phenomena.
